---
title: "Tidymodels + TidyCensus"
author:
  - name: Ethan Tenison 
    affiliation: RGK Center for Philanthropy and Community Service 
    affiliation_url: https://www.lbjdata.org/
date: "`r format(Sys.Date(), '%B %d, %Y') `"
output:
  distill::distill_article:
    code_folding: yes
    toc: yes
    toc_float: yes
    theme: theme.css
  pdf_document: default
  word_document:
    toc: yes
---


# Libraries

```{r setup, include=FALSE}

#Install all required packages 
#install.packages(c('tidyverse','tidycensus', 'tidymodels', 'vip','dotwhisker'))


#Libraries
library(tidyverse)    # data cleaning and visualization 
library(tidycensus)   # for accessing census data 
library(tidymodels)   # suite of stats modeling packages
library(vip)          # for visualizing variable importance
library(dotwhisker)   # for visualizing regression results

#Rmarkdown Options 
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# TidyCensus 

If you need to request a key go here: https://api.census.gov/data/key_signup.html

```{r pulling_data}

#Set your API KEY
#census_api_key("YOUR KEY HERE")

#View all the variables available
v19 <- load_variables(2019, "acs5", cache = TRUE)


#List out appropriate variables
median_housing_costs <- "B25105_001"
median_household_income <- "B19013_001"
avg_household_size <- "B25010_001"
housing_units <- "B25008_001"
school_age <-
  c("B14001_004", "B14001_005", "B14001_006", "B14001_007")
medicare_coverage <- "B992706_001"
hispanic <- "B03001_001"
nonfamily <- "B09020_012"


#acs variables
vars <-
  c(
    median_housing_costs,
    median_household_income,
    avg_household_size,
    housing_units,
    medicare_coverage,
    hispanic,
    "B14001_004",
    "B14001_005",
    "B14001_006",
    "B14001_007",
    nonfamily
  )

acs <- get_acs(
  geography = "tract",
  variables = vars,
  year = 2019,
  state = "TX"
)
```

# Data cleaning acs data 

```{r data_cleaning}


df <- acs |>
  pivot_wider(names_from = variable, values_from = c(estimate, moe)) |>
  select(!starts_with("moe")) |>
  rename(
    "hispanic" = "estimate_B03001_001",
    "med_hs_in" = "estimate_B19013_001",
    "hs_units" = "estimate_B25008_001",
    "hs_size" = "estimate_B25010_001",
    "med_hs_cost" = "estimate_B25105_001",
    "medicare" = "estimate_B992706_001",
    "nonfamily" = "estimate_B09020_012"
  ) |>
  mutate(
    school_age = estimate_B14001_004 + estimate_B14001_005 +
      estimate_B14001_006 + estimate_B14001_007,
    school_enrol_hshld = school_age / hs_units,
    medicare_hshld = medicare / hs_units,
    hispanic_hshld = hispanic / hs_units,
    nonfamily_hshold = nonfamily/ hs_units
  ) |>
  select(!starts_with("estimate"),
         -c(GEOID, NAME, hs_units, school_age,
            medicare, hispanic, nonfamily))

df <- na.omit(df)

```

# Splitting data 

```{r splitting}



```

# Picking the model 

hs_size ~ med_hs_in + med_hs_cost + school_enrol_hshld + medicare_hshld + hispanic_hshld

# Build model

```{r model}


```


# Fit model 

```{r fit}


```

# Tidying the results

```{r tidy}



```
# Dot and whisker plot

```{r dotwhisker}

tidy(lm_fit) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))

```
# Prediction

```{r prediction}




```

# Evaluating performance 

* Linearity
* Errors (Residuals) are normally distributed
* Homoskedastity of errors
* Independence of observations 

```{r eval}

par(mfrow = c(2, 2)) # plot all 4 plots in one plot
plot(lm_fit$fit,  # Select the fit object from the lm_fit list object 
     pch = 16,    # optional parameters to make points blue
     col = '#006EA1')


```

# Evaluating Model

```{r eval_mod}


```
# Variable Importance 

```{r vip}


```
# Check for accuracy 

```{r check accuracy}


test_results <- predict(lm_fit, new_data = test_data) |>
  bind_cols(test_data)

ggplot(data = test_results,
       mapping = aes(x = .pred, y = hs_size)) +
  geom_point(color = '#006EA1') +
  geom_abline(intercept = 0, slope = 1, color = 'orange') +
  labs(title = 'Linear Regression Results - Test Set',
       x = 'Predicted Size',
       y = 'Actual Size')

```

# Lets use Random Forest instead!

```{r rf}

```

